Заметки про квантовые группы, интегрируемые системы и все такое

### 2018-11-24 AffineLieAlgebras/Simons/paper.tex
Minor
Various changes

\begin{itemize}
\item Boltzmann weights for $A_{4}$
\item Boundary weights ?
\item Critical point, isotropicity

@@

\begin{equation}
  \label{eq:12}
  \begin{array}{l}
    W\left(                                          
        \begin{array}{cc}
          2 & 1 \\
          1 & 2
        \end{array}
      \right)=
    W\left(
        \begin{array}{cc}
          3 & 4 \\
          4 & 3
        \end{array}
      \right)=   W\left(                                         
        \begin{array}{cc}
          1 & 2 \\
          2 & 1
        \end{array}
      \right)= \\
      W\left(                                                  
        \begin{array}{cc}
          4 & 3 \\
          3 & 4
        \end{array}
      \right)=\sqrt{\frac{2}{1+\sqrt{5}}}+\sqrt{\frac{1}{2}\left(1+\sqrt{5}\right)\\
    W\left(                                      
        \begin{array}{cc}
          2 & 3 \\
          1 & 2
        \end{array}
      \right)=
    W\left(                                         
        \begin{array}{cc}
          3 & 2 \\
          4 & 3
        \end{array}
      \right)=    W\left(                                         
        \begin{array}{cc}
          3 & 4 \\
          2 & 3
        \end{array}
      \right)=    W\left(                                         
        \begin{array}{cc}
          2 & 1 \\
          3 & 2
        \end{array}
      \right)=\\
    W\left(                                       
        \begin{array}{cc}
          3 & 2 \\
          2 & 1
        \end{array}
      \right)=
    W\left(                                                       
        \begin{array}{cc}
          1 & 2 \\
          2 & 3
        \end{array}
      \right)=    W\left(                                                      
        \begin{array}{cc}
          4 & 3 \\
          3 & 2
        \end{array}
      \right)=
    W\left(                                                        
        \begin{array}{cc}
          2 & 3 \\
          3 & 4
        \end{array}
      \right)=
\sqrt[4]{\frac{2}{1+\sqrt{5}}}\\
    W\left(
        \begin{array}{cc}
          3 & 2 \\
          2 & 3
        \end{array}
      \right)=    W\left(
        \begin{array}{cc}
          2 & 3 \\
          3 & 2
        \end{array}
      \right)=2

@@

\begin{equation}
  \label{eq:6}

\end{equation}

@@

\section*{Acknowledgements}
\label{sec:acknowledgements}

This research was started during my stay at Simons center for geometry and physics in Stony Brook
university. I thank Simons foundation for the financial support during my stay, the organizers of
the program ``Conformal geometry'' and Simons center staff for their hospitality. I also thank
Christian Hagendorff(?) and Dmitry Chelkak for fruitful discussions.

This work  is supported by
the Chebyshev Laboratory (Department of Mathematics and Mechanics,
Saint-Petersburg State University) under the grant 11.G34.31.0026
of the Government of the Russian Federation and by the Dynasty foundation.

I am grateful to my wife Alexandra for her constant support and care.

>>>>

\begin{itemize}
\item Boltzmann weights for $A_{4}$
\item Boundary weights ?
\item Critical point, isotropicity
\end{itemize}

@@

\begin{equation}
  \label{eq:12}
  \begin{array}{l}
    W\left(                                          
        \begin{array}{cc}
          2 & 1 \\
          1 & 2
        \end{array}
      \right)=
    W\left(
        \begin{array}{cc}
          3 & 4 \\
          4 & 3
        \end{array}
      \right)=   W\left(                                         
        \begin{array}{cc}
          1 & 2 \\
          2 & 1
        \end{array}
      \right)= \\
      W\left(                                                  
        \begin{array}{cc}
          4 & 3 \\
          3 & 4
        \end{array}
      \right)=\sqrt{\frac{2}{1+\sqrt{5}}}+\sqrt{\frac{1}{2}\left(1+\sqrt{5}\right)}\\
    W\left(                                      
        \begin{array}{cc}
          2 & 3 \\
          1 & 2
        \end{array}
      \right)=
    W\left(                                         
        \begin{array}{cc}
          3 & 2 \\
          4 & 3
        \end{array}
      \right)=    W\left(                                         
        \begin{array}{cc}
          3 & 4 \\
          2 & 3
        \end{array}
      \right)=    W\left(                                         
        \begin{array}{cc}
          2 & 1 \\
          3 & 2
        \end{array}
      \right)=\\
    W\left(                                       
        \begin{array}{cc}
          3 & 2 \\
          2 & 1
        \end{array}
      \right)=
    W\left(                                                       
        \begin{array}{cc}
          1 & 2 \\
          2 & 3
        \end{array}
      \right)=    W\left(                                                      
        \begin{array}{cc}
          4 & 3 \\
          3 & 2
        \end{array}
      \right)=
    W\left(                                                        
        \begin{array}{cc}
          2 & 3 \\
          3 & 4
        \end{array}
      \right)=
\sqrt[4]{\frac{2}{1+\sqrt{5}}}\\
    W\left(
        \begin{array}{cc}
          3 & 2 \\
          2 & 3
        \end{array}
      \right)=    W\left(
        \begin{array}{cc}
          2 & 3 \\
          3 & 2
        \end{array}
      \right)=2

@@

\begin{equation}
  \label{eq:6}
  1
\end{equation}

@@

%%  
%%  \section*{Acknowledgements}
%%  \label{sec:acknowledgements}
%%  
%%  This research was started during my stay at Simons center for geometry and physics in Stony Brook
%%  university. I thank Simons foundation for the financial support during my stay, the organizers of
%%  the program ``Conformal geometry'' and Simons center staff for their hospitality. I also thank
%%  Christian Hagendorff(?) and Dmitry Chelkak for fruitful discussions. 
%%  
%%  This work  is supported by
%%  the Chebyshev Laboratory (Department of Mathematics and Mechanics,
%%  Saint-Petersburg State University) under the grant 11.G34.31.0026
%%  of the Government of the Russian Federation and by the Dynasty foundation. 
%%  
%%

### 2018-11-24 AffineLieAlgebras/TensorProducts/paper.tex
Various changes

%\usepackage{amsfonts}
\usepackage{pb-diagram}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage{iopams,setstack}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}

@@

\ftc{Recurrent formulae for graded tensor product multiplicities and cluster algebras}

@@

\begin{abstract}

We show the connection of formulae for tensor product multiplicities with cluster algebras. 
\noindent{\it Keywords\/}: Lie algebra, tensor product, multiplicity, cluster algebra, grading
\end{abstract}

@@

@@

 In the paper \cite{feigin1999generalized} it was shown that the module $V$ of $\gf[[t]]$ is
 generated by the action of lowering operators on the tensor product of highest weight vectors
 $v_{1}\otimes\dots \otimes v_{n}$. The algebra $\gf[[t]]$ is graded by the degree of polynomial
 and the algebra $U(\gf[[t]])$ is also graded. Then we have the induced filtration on $V$:
 \begin{equation}
   \label{eq:2}
   F^{i}V=U^{\leq i}(\gf[[t]])v_{1}\otimes\dots\otimes v_{n}.
 \end{equation}
 Note that $F^{0}V$ is the irreducible $\gf$-submodule corresponding to the highest weight in the
 tensor product decomposition
 \begin{equation}
   \label{eq:3}
   L^{\mu_{1}}\otimes\dots \otimes L^{\mu_{n}}=L^{\nu_{1}}\oplus\dots\oplus L^{\nu_{k}}
 \end{equation}

\item Fusion product of Virasoro modules, fusion in WZW models and Feigin-Loktev tensor product 
\item KR-formula
\item Weyl character formula and formulas for tensor product
\item Weyl character formula for graded tensor product
\item Computations: examples, performance
\item Cluster algebra perspective
\end{itemize}

\cite{di2013quantum} 
\bibliography{cluster}{} 
\bibliographystyle{iopart-num}

>>>>

%\usepackage{amsfonts}
\usepackage{pb-diagram}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage{iopams,setstack}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}

@@

\ftc{Feigin-Loktev fusion product and formulas for graded tensor product multiplicities}

@@

\begin{abstract}
  Feigin-Loktev fusion product appeared from the study of operator algebra of conformal field
  theory. It generalizes tensor product of representations of simple Lie algebras by introducing
  grading. Graded product multiplicities are generalized Kostka polynomials. We start with the
  fermionic formulae for these multiplicities and derive closed explicit expressions that do not
  contain summation over integer partitions. We present closed formulas for graded tensor product
  multiplicities for the fundamental representations of simple Lie algebras of small rank. These
  formulas can be regarded as q-analogue of multinomial coefficients. 
\noindent{\it Keywords\/}: Lie algebra, tensor product, multiplicity, cluster algebra, grading
\end{abstract}

@@

In certain cases fusion modules are Weyl or Demazure modules of the current algebra
\cite{chari2001weyl,chari2006weyl,chari2014modules,chari2015demazure}. Then formulas for graded
tensor product multiplicities are known (see \cite{chari2006weyl,chari2015demazure}). Some of these
cases involve non-symmetric Macdonald polynomials \cite{0803.1146v1}.

@@

 In the paper \cite{feigin1999generalized} it was shown that the module $V$ of $\gf[[t]]$ is
 generated by the action of lowering operators on the tensor product of highest weight vectors
 $v_{1}\otimes\dots \otimes v_{n}$. The algebra $\gf[[t]]$ is graded by the degree of polynomial
 and the algebra $U(\gf[[t]])$ is also graded. Then we have the induced filtration on $V$:
 \begin{equation}
   \label{eq:2}
   F^{i}V=U^{\leq i}(\gf[[t]])v_{1}\otimes\dots\otimes v_{n},\quad F^{0}V\subset F^{1}V\subset\dots
 \end{equation}

One can show that the number of graded components $F_{j}=F^{j}V/F^{j-1}V$ is finite.

 Note that $F^{0}V$ is the irreducible $\gf$-submodule corresponding to the highest weight in the
 tensor product decomposition
 \begin{equation}
   \label{eq:3}
   L^{\mu_{1}}\otimes\dots \otimes L^{\mu_{n}}=L^{\nu_{1}}\oplus\dots\oplus L^{\nu_{k}}
 \end{equation}

Having defined the filtration on $V$ one can introduce ``graded tensor product decomposition
coefficients'' or Hilbert polynomials by the formula
\begin{equation}
  \label{eq:4}
  M^{\mu_{1},\dots \mu_{n}}_{\lambda}(t)=\sum_{k\geq 0}t^{k} \mathrm{dim}\mathrm{Hom}_{\gf} (F_{k},L^{\lambda})
\end{equation}

If we set $t=1$ these coefficients coincide with tensor product decomposition coefficients:
\begin{equation}
  \label{eq:5}
  M^{\mu_{1},\dots,\mu_{n}}_{\lambda}(1)=  M^{\mu_{1},\dots,\mu_{n}}_{\lambda}
\end{equation}

 \begin{example}
   Consider tensor powers of fundamental representation of $sl_{2}$.

 \end{example}

Tensor product as affine Lie algebra module:

Klimyk rule:
\begin{equation}
  \label{eq:6}
  \mathrm{ch}L^{\mu}\cdot\mathrm{ch}L^{\nu}=\sum_{\xi\in \mbox{weights}L^{\nu}}\mathrm{ch} L^{\mu+\xi}
\end{equation}

\item Fusion product of Virasoro modules, fusion in WZW models and Feigin-Loktev tensor product

\item Use generalized Kirillov-Reshetikhin formula:\\
    Tensor product of $n_{a,j}$ modules with highest weight $j\omega_{a}$
    \begin{equation}
      \label{eq:3}
      M^{(\bf{n})}_{\lambda}(t)=\sum_{\{m_{a,j}\}}t^{\sum_{a,b,i,j}m_{a,j}C_{\gf}^{ab}\mathrm{min}(i,j)m_{b,j}}\prod_{a,j}\left[\begin{array}{c}p_{a,j}+m_{a,j}\\ m_{a,j}\end{array}\right]_{t},
      \end{equation}
      where $C_{\gf}$ -- Cartan matrix,  the sum is restricted:
      \begin{equation}
        \label{eq:6}
        \sum_{a,j}j\omega_{a}(n_{a,j}-\sum_{b}C^{ab}m_{b,j})=\lambda
      \end{equation}
      \begin{equation}
        \label{eq:7}
        p_{a,j}=\sum_{j}\mathrm{min}(i,j) n_{a,j}-\sum_{b}\mathrm{min}(i,j)m_{b,j}
      \end{equation}
     $\left[\begin{array}{c}n\\ k\end{array}\right]_{t}=\frac{(n!)_{t}}{(k!)_{t}((n-k)!)_{t}}$, $(n)_{t}=1+t+\dots t^{n-1}$.

\item Weyl character formula and formulas for tensor product
\item Weyl character formula for graded tensor product
\item Computations: examples, performance
\item Cluster algebra perspective
\end{itemize}

\section{Simple example}
\label{sec:simple-example}

Let's remind the properties of $q$-binomial coefficients
(http://mathworld.wolfram.com/q-BinomialCoefficient.html). Recurrent relation for binomial coefficients
\begin{equation}
  \label{eq:7}
  \left(\begin{array}{c}n\\ k\end{array}\right)=\left(\begin{array}{c}n-1\\
      k-1\end{array}\right)+\left(\begin{array}{c}n-1\\ k\end{array}\right) 
\end{equation}
is easy to prove:
\begin{equation}
  \label{eq:8}
  \frac{n!}{k!(n-k)!}=\frac{(n-1)!}{(k-1)!(n-k)!}+\frac{(n-1)!}{k!(n-k-1)!}
\end{equation}
Multiply by $k!(n-k)!/(n-1)!$:
\begin{equation}
  \label{eq:9}
  n=k+(n-k)
\end{equation}

Similarly for $q$-binomials we should get
\begin{equation}
  \label{eq:10}
  (n)_{q}=(k)_{q}+(n-k)_{q},
\end{equation}
but this is not the case, so we need to multiply $(n-k)_{q}$ by $q^{k}$. The resulting recurrent
relation is
\begin{equation}
  \label{eq:11}
  \left[\begin{array}{c}n\\ k\end{array}\right]_{q}=\left[\begin{array}{c}n-1\\ k-1\end{array}\right]_{q}+q^{k}\left[\begin{array}{c}n-1\\ k\end{array}\right]_{q}
\end{equation}

For $sl_{2}$ tensor product decomposition coefficients we have
\begin{equation}
  \label{eq:12}
  M^{n}_{k}(t)=\left[\begin{array}{c}n\\ k\end{array}\right]_{t}-\left[\begin{array}{c}n\\ k-1\end{array}\right]_{t}
\end{equation}
\begin{equation}
  \label{eq:13}
  M^{n}_{k}(t)=M^{n-1}_{k-1}(t)+t^{k-1}M^{n-1}_{k}+(t-1)t^{k-1}\left[\begin{array}{c}n-1\\ k\end{array}\right]_{t}
\end{equation}

$sl_2$ case here
Формула Кириллова-Решетихина дает кратность неприводимого представления со старшим весом $\lambda$ в
тензорном произведении  $n_{a,j}$ представлений со старшими весами $j\omega_{a}$, где $\omega_{a}$
-- фундаментальный вес с номером $a=1\dots \mathrm{rank}\gf$. 
\begin{equation}
  \label{eq:3}
  M^{(\bf{n})}_{\lambda}(t)=\sum_{\{m_{a,j}\}}t^{\sum_{a,b,i,j}m_{a,j}C_{\gf}^{ab}\mathrm{min}(i,j)m_{b,j}}\prod_{a,j}\left[\begin{array}{c}p_{a,j}+m_{a,j}\\ m_{a,j}\end{array}\right]_{t},
\end{equation}
здесь $C_{\gf}$ -- матрица Картана, а сумма ограничена условиями
\begin{equation}
  \label{eq:6}
  \sum_{a,j}j\omega_{a}(n_{a,j}-\sum_{b}C^{ab}m_{b,j})=\lambda
\end{equation}
\begin{equation}
  \label{eq:7}
  p_{a,i}=\sum_{j}\mathrm{min}(i,j) \left(n_{a,j}-\sum_{b}C^{ab}m_{b,j}\right)
\end{equation}
$\left[\begin{array}{c}n\\ k\end{array}\right]_{t}=\frac{(n!)_{t}}{(k!)_{t}((n-k)!)_{t}}$, $(n)_{t}=1+t+\dots t^{n-1}$.

Сперва рассмотрим простейший пример, $\gf=sl_{2}$, $t=1$. Формула Кириллова-Решетихина будет иметь вид
\begin{equation}
  \label{eq:1}
  M_{\lambda}^{(n_{1},n_{2},\dots)}=\sum_{m_{j}}\prod_{j}\left(
    \begin{array}{c}
      p_{j}+m_{j}\\
      m_{j}
    \end{array}\right),
\end{equation}
\begin{equation}
  \label{eq:2}
  \sum_{j}j\omega (n_{j}-2m_{j})=\lambda
\end{equation}
\begin{equation}
  \label{eq:4}
  p_{j}=\sum_{i}\min(i,j)(n_{i}-2m_{i})\geq 0
\end{equation}
Попробуем вычислить что-нибудь по этой формуле. Например, кратность представления со старшим весом
$1$ (индекс Дынкина) в произведении пяти фундаментальных представлений. $n_{1}=5, n_{2}=n_{3}=\dots=0$
\begin{equation}
  \label{eq:5}
  (n_{1}-2m_{1})+2(-2m_{2})+3(-m_{3})+\dots=1
\end{equation}
\begin{equation}
  \label{eq:9}
  m_{j}\geq 0
\end{equation}
\begin{equation}
  \label{eq:8}
    (5-2m_{1})+2(-2m_{2})+3(-m_{3})+\dots=1
\end{equation}
Решения: $m_{1}=2, m_{j}=0$, $m_{1}=0, m_{2}=1$. 
Для $p_{j}$ имеем в первом случае $p_{1}=1$, во втором $p_{1}=5-2=3$,$p_{2}= 5+2\cdot (-2)=1$ в итоге кратность
\begin{equation}
  \label{eq:10}
   M_{1}^{(5)}=\left(
    \begin{array}{c}
      1+2\\
      2
    \end{array}\right)+
\left(\begin{array}{c}
      1+1\\
      1
    \end{array}\right)=5
\end{equation}

\section{Proof}
\label{sec:proof}
$sl_{3}, so_{5}$ here
  \begin{itemize}
  \item $t$-deformation for the tensor product formula for $\mathfrak{sl}_{2}$:
    \begin{equation}
      \label{eq:8}
      m^{\otimes n}_{l}(t)=\left[\begin{array}{c}n-l\\n\end{array}\right]_{t}-      \left[\begin{array}{c}n-l-1\\n\end{array}\right]_{t}
    \end{equation}
  \item  $t$-deformation for the tensor product formula for $\mathfrak{sl}_{3}$:
    \begin{equation}
      \label{eq:9}
      \begin{array}{l}
      M^{n}_{p_{1},p_{2}}(t)=(n,p_{1},p_{2})_{t}-(n,p_{1}-1,p_{2})_{t}+(n,p_{1}-2,p_{2}-1)_{t}\\-(n,p_{1},p_{2}-1)_{t}+(n,p_{1}-1,p_{2}-2)_{t}-(n,p_{1}-2,p_{2}-2)_{t},        
      \end{array}
    \end{equation}
    where $(n,p_{1},p_{2})_{t}=\frac{(n!)_{t}}{(p_{2}!)_{t}((n-p_{1})!)_{t}((p_{1}-p_{2})!)_{t}}$
  \end{itemize}

Compare to Kirillov-Reshetikhin formula:

Рассмотрим теперь алгебру $sl_{3}$. Матрица Картана
\begin{equation}
  \label{eq:11}
  C=\left(\begin{array}{cc}
      2 & -1\\
      -1 & 2
    \end{array}\right)
\end{equation}
Вычислим кратность представления со старшим весом с индексами Дынкина $(1,2)$ в произведении пяти
фундаментальных представлений $L^{\omega_{1}}$. $n_{1,1}=5, n_{ij}=0$. Уравнения на $m_{aj}$:
\begin{eqnarray}
  \label{eq:12}
  (5-2m_{1,1}+m_{2,1})+2(-2m_{12}+m_{22})+3(-2m_{13}+m_{23})+4(-2m_{14}+m_{24})+\dots=1\\
  (m_{1,1}-2m_{2,1})+2(m_{12}-2m_{22})+3(m_{13}-2m_{23})+4(m_{14}-2m_{24})+\dots=2
\end{eqnarray}
сложим:
\begin{equation}
  \label{eq:13}
  (5-m_{1,1}-m_{2,1})-2(m_{12}+m_{22})-3(m_{13}+m_{23})-4(m_{14}+m_{24})+\dots=3
\end{equation}
или
\begin{equation}
  \label{eq:14}
    -(m_{1,1}+m_{2,1})-2(m_{12}+m_{22})-3(m_{13}+m_{23})-4(m_{14}+m_{24})+\dots=-2
\end{equation}
решения:
$m_{11}=2, m_{ij}=0$, $m_{12}=1, m_{ij}=0$
Вычислим $p_{ij}$. 
\begin{eqnarray}
  \label{eq:15}
  p_{11}=(n_{11}-2m_{11}+m_{21})+(-2m_{12}+m_{22})+\dots\\
  p_{21}=(m_{11}-2m_{21})+(m_{12}-2m_{22})+\dots\\
  p_{12}=(n_{11}-2m_{11}+m_{21})+2(-2m_{12}+m_{22})+\dots\\
  p_{22}=(m_{11}-2m_{21})+2(m_{12}-2m_{22})+\dots
\end{eqnarray}
В первом случае $p_{11}=1$, а остальные не нужны. Во втором $p_{11}=3, p_{21}=1, p_{12}=1,
p_{22}=2$, из которых нужно только $p_{12}$. Подставляем в формулу Кириллова-Решетихина:
\begin{equation}
  \label{eq:16}
    M_{(2,1)}^{(5)}=\left(
    \begin{array}{c}
      1+2\\
      2
    \end{array}\right)+
\left(\begin{array}{c}
      1+1\\
      1
    \end{array}\right)=5
\end{equation}
Что совпадает с разложением тензорного произведения другими способами.

Заметим, что здесь, как и в случае $sl_{2}$, при поиске решений надо учитывать те $m_{ij}$, для
которых $n_{ij}=0$.

\section*{Conlusion}
\label{sec:conlusion}

\begin{equation}
  \label{eq:17}
  \sum_{\nu=j}^{q-1} t^{(q-\nu)(q-\nu-1)} C_{q-j}^{\nu-j} C_{l-q-1}^{q-\nu-1} = 
  \sum_{\nu=j}^{q-1} t^{(q-\nu)(q-\nu-1)} C_{q-j}^{q-\nu} C_{l-q-1}^{q-\nu-1} 
\end{equation}
обозначим $q-1-\nu$ через $k$, тогда $\nu=q-1-k$, суммирование идет по $k=0\dots q-j-1$ и мы имеем
\begin{equation}
  \label{eq:18}
  \sum_{k=0}^{q-j-1} t^{(k+1)k} C_{q-j}^{k+1} C_{l-q-1}^{k} =   \sum_{k=0}^{q-j-1} t^{(k+1)k} C_{q-j}^{q-j-1-k} C_{l-q-1}^{k} = C_{l-j-1}^{q-j-1},
\end{equation}
где в последнем равенстве использовано тождество Ва
\cite{di2013quantum}

Recurrent relation for fusion modules characters is here \cite{di2015difference}
\bibliography{cluster}{} 
\bibliographystyle{iopart-num}
